\subsection{Commit}

\subsubsection*{Simple definitions}

\begin{itemize}

  \item \textbf{GROUP}: One or more planes on radar within 3nm of each other

  \item \textbf{Commit}: an informative call proword in acceptance of combat
    readiness with intent to engage, made to the Picture as a whole, regardless
    of actual targeted groups

\end{itemize}

\subsubsection*{General}

All engagements begin with a Commit informative call from the committing flight
leader, broadcast on Aux. radio.

This is a mandatory and vital call and is made when:

\begin{itemize}

  \item The Flight is committing to the picture after a picture call rather
    than a specific group

  \item The engaged fighter flight has begun to move tactically against a bogey
    group.

  \item The engaged fighter flight is forced to move tactically against a bogey
    group (aspect/range).

  \item The mission briefing dictates the flight has to move tactically against
    a bogey group.

  \item The Air Intercept Controller issues a Commit directive.

\end{itemize}

\subsubsection*{Process}

\textbf{FL RIO}: Via comms examples below they will issue a commit call when
appropriate

\subsubsection*{Examples}

\textbf{FL RIO (Aux)}: "Spectre 1, Commit, target group Alpha, BRAA,
One-Three-Zero, 50, twelve thousand."

\textit{or}

\textbf{MAGIC (Aux)}: "Spectre 1, Magic, commit group Bravo, BRAA,
One-Three-Zero, 50, twelve thousand."\\
\textbf{FL RIO (Aux)}: "Spectre 1, Commit Bravo, BRAA One-Three-Zero, 50,
twelve thousand."

\textit{or}

\textbf{MAGIC (Aux)}: "Spectre 1, Magic, PICTURE Single GROUP Bravo,
BRAA, One-Three-Zero for 50, twelve thousand, track WEST, BOGEY."\\
\textbf{FL RIO (Aux)}: "Spectre 1, Commit"

\boxed{
  Spectre commits AND targets a specific group because the group met a
  commitment pre briefed range or reason.

  AIC committed Spectre against a specific group as a directive

  Spectre commits to the picture with an informative call. The preconditions
  for this are that they heard the PICTURE from AIC and commit to it.
}
