\subsection{Correlation}

\subsubsection*{Simple Definitions}

\begin{itemize}

  \item Correlation: The verification process of cross-referencing" the same
    group: A call to reference the same group between AIC and fighters
    within 3 degrees and 3nm accuracy.

  \item Declare: Inquiry as to the ID of a correlated group, a specific track,
    or target

  \item Labelling: The process of giving aircraft groups "friendly names" eg
    "North group", "single group", "Fulcrum group", "Charlie group", "Two-ship
    group".

\end{itemize}

\subsubsection*{General}

Correlation is where the flight and the Air Intercept Controller both ensure
they are agreeing on the same Group of interest by the Flight calling a
Bearing, Range and Altitude to the group the AIC has called out, or the group
the fighter is committing on. Correlation in the real world accepts 3nm error
or 3 degrees error.

This is done with the "Declare" proword.

\boxed{
  Note: that the Tomcat accepts a degraded DL. You cannot correlate a group you
  do not have radar contact on. You must not confuse a DL only object, with a
  combined HAFU from DL and Radar.
}

\subsubsection*{Examples}

\textbf{MAGIC (Aux):} "Magic, Spectre 1, commit group Bravo, BRAA,
One-Three-Zero for 50, twelve thousand."
\\
\textbf{FL RIO (Aux):"} Spectre 1, Declare Group BRAA, One-Three-One for
fifty-one, Twelve thousand."
\\
\textbf{MAGIC (Aux):} "Magic, Spectre 1, Group Bravo, BANDIT"
\\
\textbf{FL RIO (Aux):} "Spectre 1, Commit Group Bravo"

\boxed{%
  Magic directs Spectre to commit on a group called "Bravo" with a BRAA Spectre
  RIO looks for it, finds a radar return and then correlates his BRAA with a
  contact in front of him Magic correlates the information and tells him it's
  the same Group Bravo and additionally uses amplification on the word BANDIT to
  say that the group is not HOSTILE.

  Spectre acknowledges the commit.
}
