\subsection{Employment (NLT 35nm)}

\subsubsection*{Simple Definitions}

\begin{itemize}

  \item Fox-3: Proword for declaring an Active Radar Homing Missile launch

  \item Gate: Proword for applying full Reheat or Afterburner on the throttle

  \item Buster: Proword for setting full non afterburner throttle (MILitary
    power)

  \item Gimbals: The limit of the radar Azimuth left or right (or elevation)

  \item Crank: A manouver taking the nose of the plane left or right until the
    radar contact is on the Gimbals (but still on scope)

  \item "fox-3, two ship": Indicates missiles were employed on two contacts in
    a group

  \item "second fox-3” indicates a second launch on the same contact.

\end{itemize}

\subsubsection*{General}

The RIO will call for "Go Gate!" at their discretion, in order to gain an
energy advantage.

Flight will shoot and crank together, \textbf{whether they have fired or not}.

\subsubsection*{Employ}

The actual employment geometry will depend on the altitude and aspect of the
targeted group. For head-on, co-alt engagements, the 40-35nm time slice will
consist of a desired supersonic acceleration before launch, followed by a crank
and descend. The use of Gate should not be too soon as to give away intentions
and use too much fuel, or too late as to not make use of the energy.

\subsubsection*{Crank}

The direction to Crank is often decided by safety but is a hard turn to the
radar gimbals in the direction of (in order of importance):

\begin{enumerate}
  \item Away from danger, SAM zones, other Enemy fighters
  \item Towards Safe zones or unengaged friendly fighters
  \item Cold side of the bandit
\end{enumerate}


\subsubsection*{Example}

\textbf{FL RIO (Pri):} "Spectre 1, Go Gate!" \\
\textbf{FL RIO (ICS):} "Steady up! Centre the T!" \\
\textbf{FL RIO (Aux):} "Spectre 1-1, Fox three, Two-Ship,
\textit{<Bulls>}, One two zero,
thirty-five!" \\
\textbf{FL RIO (Pri):} "Crank left! Go Idle! Ten degrees nose down!" \\
\textbf{WM RIO (Aux):} "Spectre 1-2, Fox three, Two-Ship, same group"

\boxed{%
  The FL RIO commands an (ideally supersonic) acceleration until the shot
  parameters to be met and tells the pilot to centre the T and steady up. Once
  established, shoots two fox 3 at two targets in the group, and announces it
  on Broadcast Aux in Bulleye format for  maximum situational awareness for
  friendly forces. WM RIO announces his shots and FL RIO commands crank and
  descent on Pri to reduce closure rate.

  There are now 4 missiles, targeting two contacts in flight.
}

