\subsection{Identification \& Classification}

\subsubsection*{Simple Definitions}

\begin{itemize}
  \item ID: Directive to identify the target. Once assessed, return a
    classification.

  \item Classification: The type of ID

  \begin{itemize}
    \item Bogey - unknown (and no checks done)
    \item Friendly - IFF positive
    \item Bandit: ID fails IFF response (on AWG-9 and no implication to fire)
    \item Hostile: ID that a group can be fired upon
  \end{itemize}

  \item Declare: Brevity Code Inquiry as to the ID of a specified track(s),
    target(s), or correlated GROUP. Reply with a Classification and Position.

  \item VID ("vee-i-dee"): Directive to visually identify a target for aircraft
    type and operator markings / registration"

\end{itemize}

\subsubsection*{General}

Classification is a process which is continuously checked during the
intercept. Classification seeks to determine (by anyone in the Area of
Operations) if a Group on Radar is Friendly or should be engaged.

\boxed{
  Note: Between Bogey and Hostile are a range of classifications that will be
  dealt with under CT, such as \textbf{SPADES, RIDER, SQUAWKING, GOPHER,
  OUTLAW} to describe various types of partially identified aircraft groups.
  These prowords explain what the ID is, e.g. origin - \textbf{Outlaw}, lack of
  IFF - \textbf{Spades}, unknown but friendly IFF \textbf{Gopher}, unknown but
  flying in a friendly corridor - \textbf{Rider}.
}

\textbf{Identification must be done prior to weapons employment}. It requires
\textbf{two} factor checks. The most common in DCS is the factor of Coalition
ID which wraps up IFF and hostile behaviour in one neat package of "Blue or
Red". From a LotATC point of view, they do not necessarily have a blue or red
simplicity, so often the Fighter is more knowledgeable than AIC with regards to
IFF status and can classify as BANDIT themselves.

Identification will continue. A second form of classification \textbf{must}
occur apart from IFF. This is most commonly done by intentions but can include
Visual:

\begin{itemize}
  \item Spiked longer than an IFF is needed
  \item Shoots
  \item Maneuvers aggressively in aspect or towards a friendly
  \item Visually Identified (VID) for example by Television Camera System
\end{itemize}

The Flight must conclude HOSTILE before weapons employment and especially at or
before a favourable employment on the \textbf{Timeline}. Any contributing
agency or flight can make a HOSTILE ID.

\boxed{
  Note: RIO's should avoid using the HOSTILE designation on AWG-9 'CAP' for
  anything not complying to HOSTILE intentions, i.e. it is not a case of going
  through checking IFF and updating as HOSTILE directly. That is not how
  classification works.
}
