\subsection{Introduction}

Whilst there are some overlaps with the A-A TTP's, this chapter handles the
\textbf{very specific SOP of Phoenix} deployment on Timeline and omits some of
the deeper detail in the A-A TTP 1 and 2. The Phoenix deployment is lost in
1990's process, whilst the AMRAAM took over as the primary air to air weapon.
Some elements of the modern process have been converted to 108\th usage.
Additionally, we make special mention of Crew interactions. Both Pilot and RIO
need to be fairly confident in what the Timeline entails, although absolute
proficiency at MQT stage will be difficult and not expected until CT. This
topic is overloaded, but the required context means that the detail and outline
must be complete in order to be read and understood. It is Tomcat specific for
Timeline due to ranges.

The purpose of the detail in this section is to provide a "\textbf{default shot
parameter}" for the bread and butter weapon of the F-14B Tomcat. For other
weapons such as AIM-7, AIM-9, the briefing should cover any important
differences in Ranges.

The Mission briefing will always cover the Commit and Engagement
\textbf{ranges}, because the Phoenix can be fired at much further ranges. The
\textbf{NLT 35nm shot} is an "aware fighter" range, that is, it is a shot range
designed to kill a manoeuvring fighter, rather than force an aircraft to defend
or target bombers. 35nm is not always the best range for a 'mission'.

More distant ranges can be used in missions where delay and follow up, or
skate/aborting is part of the mission. Long shots are part of CT/Briefing. STT
shots are part of CT/Briefing

\textbf{The primary radar mode to employ Phoenix from is TWS} with the RIO's
sorting as per 108\th Contract SOP: Lead has "\textbf{Left/Lead/High}" and WM
has "\textbf{Right/Trail/Low}" by selecting the primary contact with one shot
each. This is for maximum surprise element on the timing of the shot to the
bandit. TWS is both effective and flexible. STT shooting is advised for Long
Range intercepts and fleet defence which is currently outside scope and a
briefed item.

'Doubling shots' to increase PK is a Flight Lead choice to be briefed and is
not default. This is not PvP. PvP "pk increases" will be dealt with in A-A TTP
2.

Employment sorting on Heavy groups is a \textbf{mandatory} briefed item.
